### 2016-11-30 SimulatedData.Rmd
revising simulation function

---
title: "Trait-matching simulation" 
author: "Ben Weinstein"
output:
  html_document:
    toc: true
    number_sectionsf: true
    theme: spacelab
---

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(pez)
library(ggplot2)
library(knitr)
library(dplyr)
library(stringr)
opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=4,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")
```

#Simulation

What are the data we want to simulate?

size_x = Number of level 1 (eg. Plants)
size_y = Number of level 2 (eg. Pollinators)
traits_x = Trait frame for level 1
traits_y = Trait frame for level 2
phylogeny_x = phylogeny for level 1
phylogeny_y = phylogeny for level 2
Obs = a x by y matrix of visits

```{r}
#sim_assemblages<-function(x=,y=,traits=
## Process Model
```

```{r,fig.height=5,fig.width=8,eval=T}
#Number of consumer species
h_species=10
plant_species=10
cameras<-5
days<-3

#Bill sizes
Bill<-rpois(h_species,10)

#Corolla sizes
Corolla<-rpois(plant_species,15)

#Subtract both and take absolute value, convert cm
traitmatch<-abs(sapply(Corolla,function(x) x - Bill)/10)

#regression slopes
#traits
beta1_mu<- -1

#abundance

#species variance in slopes
beta1_sigma<- 0.2

#Species alpha_mu 
alpha_mu<- 3
alpha_sigma<- 0.2

#Poisson overdipersion
tauE<-10

#species level
detection= inv.logit(rnorm(h_species,0,1.67))
beta1<-rnorm(h_species,beta1_mu,beta1_sigma)
alpha<-rnorm(h_species,alpha_mu,alpha_sigma)
```

#Compute true interaction matrices

```{r,eval=T}
#for each species loop through and create a replicate dataframe
obs<-array(dim=c(h_species,plant_species,cameras,days))
lambda<-array(dim=c(h_species,plant_species,cameras))
N<-array(dim=c(h_species,plant_species,cameras))

#create intensities
for(x in 1:h_species){
  for (y in 1:plant_species){
    for (z in 1:cameras){
      lambda[x,y,z]<-exp(alpha[x] + beta1[x] * traitmatch[x,y] + rnorm(1,0,1/sqrt(tauE)))
  }
  }
}

#draw latent states
for(x in 1:h_species){
  for (y in 1:plant_species){
    for (z in 1:cameras){
      # true latent count
      N[x,y,z]<-rpois(1,lambda[x,y,z])
    }
  }
}
```

##View correlation in simulated latent state

```{r}
mdat<-melt(N)
colnames(mdat)<-c("Bird","Plant","Camera","Interactions")

traitmelt<-melt(traitmatch)
colnames(traitmelt)<-c("Bird","Plant","traitmatch")

mdat<-merge(mdat,traitmelt,c("Bird","Plant"))
ggplot(mdat,aes(x=traitmatch,y=Interactions,col=as.factor(Bird))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args = list(family = "poisson")) + labs(col="Bird") + xlab("Absolute value of Bill Length - Corolla Length ")
```

##View Detection Rates

```{r}
obs.state<-melt(obs)
colnames(obs.state)<-c("Bird","Plant","Camera","Day","Yobs")
obs.state<-merge(mdat,obs.state,by=c("Bird","Plant","Camera"))
ggplot(obs.state,aes(x=Interactions,y=Yobs,col=Camera)) + geom_point() + theme_bw() + geom_abline() + coord_equal()
```