### 2016-11-30 index.Rmd
add a makefile for R users that do not use Rstudio

---
title: "Inferring ecological interactions"
---

#Aims

To compare methods for inferring species interactions

#Where & When

Sherbrooke, Quebec

Nov 30th - Dec 2nd, 2016

#Workshop Participants

[Dominique Gravel](), [Tim Poisot](), [Kevin Cazelles,](http://kevincazelles.fr/) [Ben Weinstein](benweinstein.weebly.com), [Ignacio Bartomeus](),Ignacio Morales Castilla, David Beauchesne, Mickaël Hedde, Mathilde Besson, Steve Vissault, Guillaume Blanchet

# Access

This website and associated analysis can be cloned

```{}
git clone https://github.com/bw4sz/Montreal_Workshop.git
```