### 2016-11-30 index.Rmd
add spreadsheet and google docs links

---
title: "Inferring ecological interactions"
---

#Aims

To compare methods for inferring species interactions

#Where & When

Sherbrooke, Quebec

Nov 30th - Dec 2nd, 2016

#Workshop Participants

[Dominique Gravel](), [Tim Poisot](), [Kevin Cazelles,](http://kevincazelles.fr/) [Ben Weinstein](benweinstein.weebly.com), [Ignacio Bartomeus](),Ignacio Morales Castilla, David Beauchesne, Mickaël Hedde, Mathilde Besson, [Steve Vissault](http://steveviss.github.io/), Guillaume Blanchet

# Access

### Website

This website and associated analysis can be cloned

```{}
git clone https://github.com/bw4sz/Montreal_Workshop.git
```

### Google doc:

[Here](https://docs.google.com/document/d/1x1YHzF-vZXbrQUIF97TNqyIJ6bXyaEiFjMsi18nJDag)

### Spreadsheet :

[Here](https://docs.google.com/spreadsheets/d/1Nps6BsDrkWMbvLa_lsgc5MT8RBtTeFX6hWlUjsRlkO0/edit?ts=583f2190#gid=1879087403)