### 2016-12-01 Data/data_explnation.Rmd
Add data (not completelly ready, but usefl)

---
title: "Data description"
author: "I. Bartomeus"
date: "12/1/2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

__Note:__ This data needs more curation and can be used to play with methods, but needs taxonomic unification, better traits, etc... before looking at results.

Data collected in 2015 across 16 sites. Each site was visited 7 times. Each time we recorded three data types:

- All interactions in 100 m transects
- Focal interactions to selected plants
- Oportunistic new interactions not detected by the last two methods (`out`)

there are three datasets so far:

__Interactions:__ This include `transect` and `visitation` data. Data can be used pooled (1445 interactions (5643 visits) general question abouts trait matching) or splited by site (`Site_ID`) and `Round`. Pollinator traits in this table are measured per individual.
```{r}
int <- read.csv("interactions.csv", h = T)
head(int)
nrow(int)
sum(int$Frequency, na.rm = TRUE)
```
Challenges:  
- long tail of singletones  
- not all species are taxonomically resolved. What to do with specimens we ony know Genus?  
- Low sample size when you split per site and round

__pollinator traits:__
```{r}
pol_tr <- read.csv("traits_pollinators.csv", h = T)
head(pol_tr)
```
Challenges:  
- We are not sure which trait are driving interactions (quite generalized system)  
- Need to inegrate traits across different orders

__plant traits:__ Not ready yet

__independent plant flower abundance__: Ca be combined by Site_ID and Round
```{r}
flw <- read.csv("flowers.csv", h = T)
head(flw)
```

__Phylogenies__ work in progress.  
- Plants: Phylomatic  
- Pollinators: Available at genus level only