### 2016-12-02 ms/intro_drafted.Rmd
First attempt to draft an introduction. Requires much much more work.

Making predictions of All Links In Ecological Networks

Introduction

Biotic interactions form the backbone of biodiversity (Bascompte & Jordano 2007). They pose selective pressures that are key to evolution (ref) and operate together with abiotic factors in constraining the distribution of biodiversity (Soberon et al. 2007). Documenting and analyzing ecological networks is essential to: (i) understanding the rules underlying pairwise species interactions (Bascompte et al. 2006), (ii) quantifying ecosystem functioning and services (Thomson et al 2012), or (iii) predicting changes in from of global change scenarios (ref). In spite of increasing efforts, it is unlikely that we will ever manage to empirically document the massive amount of interactions existing in nature (Morales-Castilla et al. 2015). Because of this, complementing field work with modelling is gaining acceptance among network ecologists (ref). In a context where novel ecosystems arise as a result of species invasions or due to geographic shifts tracking climate change, our ability to predict future interactions will limit how efficiently we anticipate and respond to threats on biodiversity (i.e. species local extinctions) and on humans (i.e. spread of infectious disease or plagues affecting to crops).

Species characteristics or traits are at the core of biological interactions (Bartomeus et al. 2016). This is because traits are the results of species adaptations to exploit abiotic or biotic resources. In fact, traits are rooted in all the principal theoretical mechanism proposed to explain when two species can interact, and hence are the perfect candidates to become the backbone of any framework that aims to predict species interactions. First, the idea of morphological trait matching directly relates to the idea that two species can only interact if there is certain concordance between the traits of both partners. For example, in fishes predator mouth gap  has to be bigger than prey body size in order to get a predatory interaction (Gravel et al. 2013). This idea is formalized in the niche model theory, which in its simpler form postulates that [species optimum is placed along a niche axes?…] (Williams...) and has been subsequently refined to include [more axes, prob] … (Gravel et al. 2013, Bartomeus et al. 2016). Second, species traits also determine the range of environmental conditions on which species can survive and co-occurrence in space and time is precisely the first requisite for two species to interact (Morales-Castilla et al 2015). Hence, patterns emerging through processes governed by island biogeography theory can be also viewed at the light of traits. Third, traits are the result of evolutionary history. Phylogenetic information can constrain trait evolution either directly (affecting upon trait matching) or indirectly (via niche conservationism, for example). Finally, it’s worth noting that traits are not only related to the niche differences among species, but also to the fitness effects gained by both interaction partners. This can potentially establish a link with to modern coexistence theory. In fact, neutral expectations postulating that species abundances are the main drivers of interactions, are not incompatible with the underlying role of species traits shaping species abundances via effects in fitness (Bartomeus et al. 2016).

We have showed that traits are a common currency that link mechanistic understanding of species interactions, but given that different traits are involved in different processes (e.g. traits involved in trait matching are not the same involved in species co-occurrence), one may think using traits for inferring biotic interactions would rapidly become intractable. Luckily, recent research has shown that traits can be organized in ecological syndromes and that less than five trait axis may suffice to describe ecological interactions (Eklöf).

Indeed, on the quest for describing species interactions, traits had played a pivotal role and several methods have recently emerged to test hypotheses of trait matching for pairwise interactions (Stang et al 2006, Delhing et al. 2014, Bartomeus et al 2016) or understand the phylogenetic structure of ecological networks (Rezende et al. 2007). However, generating biological meaningful predictions of species interactions and its emerging properties have proven challenging (Olito and Fox 2014). An ideal method is grounded in ecological theory,  which makes the description of ecological interactions interpretable. This method should be able to not only describe, but also predict ecological interactions based on a set of simple parameters. Importantly, this predictions should be accompanied by an associated uncertainty. In addition, given the difficulty of empirically assessing complete interaction networks, this method should be also robust to the use of incomplete data.

Here, we review the main methods proposed to evaluate plant-pollinator interactions with emphasis on its utility for predicting the position of a new species in the network. We use simulated and real data … Our objectives is to provide a guideline of the strengths and limitations of different methods in relation to the type of question pursued and data availability. There is an urgent need to better understand how communities interact and predict new interactions which will inevitably occur as a result of human induced rapid environmental change. In addition, our understanding of how species interact is limited to the few interactions are directly observable such as pollination of predation, but key species interactions not directly observable remains unlocked. Interactions inferences may be the only option for describing and predicting interactions in soil food webs or bacteria-phage networks.