### 2016-12-07 SimulatedData.Rmd
link to manuscript

---
title: "Trait-matching simulation" 
author: "Ben Weinstein"
output:
  html_document:
    toc: true
    number_sectionsf: true
    theme: spacelab
---

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(bipartite)
library(ggplot2)
library(knitr)
library(boot)
library(tidyr)
library(dplyr)
library(stringr)
library(R2jags)
opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=5,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")

#source helper function
source('Bayesian/Functions.R')
```

#Simulation

What are the data we want to simulate?

size_x = Number of level 1 
size_y = Number of level 2
traits_x = Trait frame for level 1
traits_y = Trait frame for level 2
Obs = x by y matrix of visits

```{r}
simA<-function(size_x,size_y,traits_x,traits_y,beta1_sigma= 0.4,alpha_sigma= 0){

  #Subtract both and take absolute value, convert cm
  traitmatch<-abs(sapply(traits_y,function(x) x - traits_x))

  #regression slope
  #traits
  beta1_mu<- -1

  #species variance in slopes
  beta1_sigma<- 0.1

  #Species alpha_mu 
  alpha_mu<- 3
  alpha_sigma<- 0

  #species level
  beta1<-rnorm(size_x,beta1_mu,beta1_sigma)
  alpha<-rnorm(size_x,alpha_mu,alpha_sigma)

  #for each species loop through and create a replicate dataframe
  obs<-array(dim=c(size_x,size_y))
  N<-array(dim=c(size_x,size_y))

  #create intensities
  for(x in 1:size_x){
    for (y in 1:size_y){

        #intensity
        N[x,y]<-inv.logit(alpha[x] + beta1[x] * traitmatch[x,y])

        #draw one state
        obs[x,y]<-rbinom(1,1,N[x,y])
    }
  }

#draw intensity

tx<-data.frame(I=1:length(traits_x),TraitI=traits_x)
ty<-data.frame(J=1:length(traits_y),TraitJ=traits_y)

#view trait matching
dat<-melt(obs)
colnames(dat)<-c("I","J","Interactions")

dat<-merge(dat,tx)
dat<-merge(dat,ty)

#define data types, make species
dat$I<-letters[dat$I]
dat$J<-letters[dat$J]

return(dat)
}
```

```{r}
size_x = 5
size_y = 19

#Bill sizes
traits_x<-rpois(size_x,10)

#Corolla sizes
traits_y<-rpois(size_y,10)

dat<-simA(size_x=size_x,size_y=size_y,traits_x=traits_x,traits_y=traits_y)

#view heatmap
heatmap<-function(dat){
  ggplot(data=dat,aes(x=I,y=J,fill=Interactions)) + geom_tile() +    scale_fill_continuous(low="white",high="red")
}

heatmap(dat)

traitplot<-function(dat){
    ggplot(data=dat,aes(y=Interactions,x=abs(TraitI-TraitJ))) + geom_point() + geom_smooth(aes(group=1),method="glm",method.args=list(family="binomial")) + labs(x="Difference in traits",y="Probability of Interaction")
}

traitplot(dat) + theme_bw()
```

# Intercept

```{r,eval=T}
fitb<-function(dat,algorithm="Intercept"){

  #format traitmatch as matrix
  dat$Traitmatch<-abs(dat$TraitI-dat$TraitJ)
  Traitmatch<-acast(data=dat,I~J,value.var="Traitmatch")

  if(algorithm == "Intercept"){

  runs<-10000

  #Source model
  source("Bayesian/Intercept.R")

  #print model
  print.noquote(readLines("Bayesian//Intercept.R"))

    #for parallel run
    Yobs=dat$Interactions
    Bird=as.numeric(as.factor(dat$I))
    Plant=as.numeric(as.factor(dat$J))
    Birds=length(unique(dat$I))
    Traitmatch=Traitmatch
    Plants=length(unique(dat$J))
    Nobs<-length(Yobs)

    #MCMC options
    ni <- runs  # number of draws from the posterior
    nt <- 1  #thinning rate
    nb <- max(0,runs-500) # number to discard for burn-in
    nc <- 2  # number of chains

    modelDat<-list("Yobs","Bird","Plant","Plants","Birds","Nobs")

    ParsStage <- c("alpha","alpha_mu","ynew","fit","fitnew")

m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Intercept.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))
} else

    if(algorithm=="Binomial") {

  runs<-10000

  #Source model
  source("Bayesian/Binomial.R")

  #print model
  print.noquote(readLines("Bayesian//Binomial.R"))

    #for parallel run
    Yobs=dat$Interactions
    Bird=as.numeric(as.factor(dat$I))
    Plant=as.numeric(as.factor(dat$J))
    Birds=length(unique(dat$I))
    Traitmatch=Traitmatch
    Plants=length(unique(dat$J))
    Nobs<-length(Yobs)

    #Parameters to track
    ParsStage <- c("alpha","beta","alpha_mu","beta_mu","ynew","fit","fitnew")

    #MCMC options
    ni <- runs  # number of draws from the posterior
    nt <- 1  #thinning rate
    nb <- max(0,runs-500) # number to discard for burn-in
    nc <- 2  # number of chains

    modelDat<-list("Yobs","Bird","Plant","Plants","Traitmatch","Birds","Nobs")

m1<-do.call(jags.parallel,list(data=modelDat,parameters.to.save=ParsStage,model.file="Bayesian/Binomial.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc,DIC=F))  
            }

    m1$Algorithm<-algorithm
    return(m1)
}
```

# Interecpt Model
```{r}
intercept<-fitb(dat,"Intercept")
```

# Binomial Regression
```{r}
binomial<-fitb(dat,"Binomial")
```

#Extract parameters from the model

```{r}
getpars<-function(m1,n.iter=100,nt=1,nb=0,pars=c("beta1","alpha1")){  
  parsm1<-extract_par(m1)
  parsm1$Model<-m1$algorithm
  return(parsm1)
}
```

```{r}
#bind them together
datpars<-list(intercept,binomial)

datpars<-lapply(datpars,getpars)
```

#Predict new species

```{r}
size_x = 2
size_y = 5

#Bill sizes
traits_x<-rpois(size_x,10)

#Corolla sizes
traits_y<-rpois(size_y,10)

newtraits<-simA(size_x=size_x,size_y=size_y,traits_x=traits_x,traits_y=traits_y)

#calculate traitmatch
newtraits$Traitmatch<-abs(newtraits$TraitJ-newtraits$TraitI)

predictM<-function(m1,newtraits){
  parsm1<-extract_par(m1)

  #matching function
a<-parsm1 %>% filter(par %in% c("alpha_mu","beta_mu")) %>% select(Draw,Chain,par,estimate) %>% dcast(.,Draw + Chain ~ par,value.var="estimate") %>% group_by(Draw,Chain)%>%  do((pred.binomial(alpha=.$alpha_mu,beta=.$beta_mu,newdata=newtraits))) %>% group_by(I,J) %>% summarize(mean=mean(value),lower=quantile(value,0.95),higher=quantile(value,0.05))

return(a)
}

predt<-predictM(binomial,newtraits)
```

# Model evaluation

```{r}
parsall<-rbind_all(list(parsm1,parsm2))
```

## Posterior predictive check

```{r,fig.height=4,fig.width=8}
fitstat<-droplevels(parsall[parsall$par %in% c("fit","fitnew"),])
fitstat<-dcast(fitstat,Draw+Model+Chain~par,value.var="estimate")

#add 1:1 line
ymin<-round(min(c(fitstat$fit,fitstat$fitnew)))
ymax<-round(max(c(fitstat$fit,fitstat$fitnew)))
ab<-data.frame(x=ymin:ymax,y=ymin:ymax)

ggplot(data=fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + coord_equal() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")
```

## Proportion of predicted links

```{r}
propc<-function(x,orig=dat){
  predweb<-acast(data=x,I~J,value.var="Index")

  #calculate network statistic
  nstat<-networklevel(predweb,index=c("connectance","nestedness"))
  data.frame(Metric=names(nstat),nstat)
}

prop_links<-parsall %>% filter(par=="ynew") %>% group_by(Model,Draw,Chain) %>% inner_join(dat) %>% do(propc(.))

```

###View trait-matching

```{r}
#merge the dataframes.
dat$Index<-1:nrow(dat)

#posterior distribution for each index
ypred<-pars %>% filter(par=="ynew") %>% group_by(Index) %>% summarize(mean=mean(estimate),lower=quantile(estimate,0.05),higher=quantile(estimate,0.95))

datpred<-merge(dat,ypred)

ggplot(datpred) + geom_point(aes(x=Traitmatch,y=Interactions),col="black") + geom_ribbon(aes(x=Traitmatch,ymin=higher,ymax=lower),alpha=0.6) + theme_bw()

ggplot(datpred) + geom_point(aes(x=Traitmatch,y=Interactions),col="black") + geom_ribbon(aes(x=Traitmatch,ymin=higher,ymax=lower),alpha=0.6) + theme_bw() + facet_wrap(~I)
```

### Compute emergent statistics

```{r}
makeN<-function(x){
  predweb<-acast(data=x,I~J,value.var="estimate")
  #calculate network statistic
  nstat<-networklevel(predweb,index=c("connectance","nestedness"))
  data.frame(Metric=names(nstat),nstat)
}

#calculate network stats for each model
nstats<-pars %>% filter(par=="ynew") %>% group_by(Draw,Chain) %>% inner_join(dat) %>% do(makeN(.))

#calculate true network stat
tweb<-acast(dat,I~J,value.var="Interactions")
tstat<-networklevel(tweb,index=c("connectance","nestedness"))
tstat<-data.frame(Metric=names(tstat),tstat)

ggplot(nstats[nstat$Metric %in% "connectance",],aes(x=nstat)) + geom_density(alpha=0.6,fill="black") + facet_wrap(~Metric,scales='free',nrow=1) + theme_bw()  + labs(x="value")
+ geom_vline(col="red",linetype="dashed",data=tstat,aes(xintercept=tstat)) 
```