### 2016-12-19 Data/data_explnation.Rmd
Add clean data

---
title: "Data description"
author: "I. Bartomeus"
date: "12/1/2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data collected in 2015 across 16 sites. Each site was visited 7 times. Each time we recorded three data types:

- All interactions in 100 m transects (`transects`)
- Focal interactions to selected plants (`focal`)
- Oportunistic new interactions not detected by the last two methods (`out`)

there are three datasets so far:

__Interactions:__ This include `transect`, `visitation` and `out` data. Data can be used pooled (2067 interactions (5732 visits) general question abouts trait matching) or splited by site (`Site_ID`) and `Round`. 
```{r}
int <- read.csv("interactions.csv", h = T)
head(int)
nrow(int)
sum(int$Frequency, na.rm = TRUE)
```
Challenges:  
- long tail of singletones  
- not all species are taxonomically resolved. What to do with specimens we ony know Genus?  
- Low sample size when you split per site and round

__pollinator traits:__
```{r}
pol_tr <- read.csv("traits_pollinators.csv", h = T)
head(pol_tr)
```
__NOTE:__ Quite a lot of NA's still.

Challenges:  
- We are not sure which trait are driving interactions (quite generalized system)  
- Need to inegrate traits across different orders (e.g. tongue lengths can be set to zero for coleoptera)

__plant traits:__ 
```{r}
plant_tr <- read.csv("traits_plants.csv", h = T)
head(plant_tr)
```
Those are complete.

__independent plant flower abundance__: Ca be combined by Site_ID and Round
```{r}
flw <- read.csv("flowers.csv", h = T)
head(flw)
```

__Phylogenies__ Need to do.
- Plants: Phylomatic? 
- Pollinators: Bees available at genus level only. For other orders I have no idea.

# combining data

Question: should we use the pooled data for the package?
```{r}
head(dat)
temp <- merge(int, pol_tr, by.x = "Pollinator_gen_sp", by.y = "Pollinator_gen_sp", all.x = TRUE)
temp2 <- merge(temp, plant_tr, by.x = "Plant_gen_sp", by.y = "Plant_gen_sp", all.x = TRUE)
pooled_data <- temp2
head(pooled_data)
```