### 2017-01-28 ms/tables/comparison.tex

\begin{sidewaystable}
\centering
\caption{Methods compared and how they apply}
\label{comparisonTab}
\begin{tabular}{llccccccccccc}
                                                & \multicolumn{5}{c}{Input Data}                                                                                                                                         & \multicolumn{1}{c}{}              & \multicolumn{1}{c}{}              & \multicolumn{1}{c}{}           & \multicolumn{1}{c}{}       & \multicolumn{1}{c}{}                   & \multicolumn{1}{c}{}                                                                                                                     \\
                                                & \multicolumn{1}{c}{abundance} & \multicolumn{1}{c}{Indirect  knowledge} & \multicolumn{1}{c}{Co-occurence} & \multicolumn{1}{c}{Trait} & \multicolumn{1}{c}{Phylogeny} & \multicolumn{1}{c}{Theory-driven} & \multicolumn{1}{c}{Type of links} & \multicolumn{1}{c}{Predictive} & \multicolumn{1}{c}{Output} & \multicolumn{1}{c}{Hypothesis testing} & \multicolumn{1}{c}{Example}                                                                                                              \\
Fourth-corner                                   & x                             &                                         &                                  & x                         &                               & yes                               & quantitative                      & mecanistic                     &                            &                                        & \begin{tabular}[c]{@{}l@{}}I\_eat (Beauschene et al. 2016)\\ Markov logic, Intuitive Logic Programming (Vacher et al. 2016)\end{tabular} \\
Maching learning                                & x                             & x                                       & (x)                              & (x)                       & (x)                           & no                                & quantitative                      & data-based                     &                            &                                        & Rohr et al. 2014, Rohr et al. 2016                                                                                                       \\
Direct mathcing-centrality                      &                               &                                         &                                  & x                         & x                             & yes                               & quantitative                      & mecanistic                     &                            &                                        &                                                                                                                                          \\
Indirect matching-centrality                    &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Niche model                                     &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Hierarchical model                              &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Null model                                      &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Joint Spatial Division and Multiplexing (JSDM)  &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Intercept                                       &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
GLM/cooccur                                     &                               &                                         &                                  &                           &                               &                                   &                                   &                                &                            &                                        &                                                                                                                                          \\
Grouping species into nodes (functional groups) &                               & x                                       &                                  & x                         &                               & no                                & Binary (quantitative)             & no                             &                            &                                        &                                                                                                                                          \\
Expert-based assesment of links                 &                               & x                                       &                                  & x                         &                               & no                                & Binary (quantitative)             & no                             &                            &                                        &
\end{tabular}
\end{sidewaystable}