### 2017-02-14 Data/data_explnation.Rmd
Data completely clean and complete.

---
title: "Data description"
author: "I. Bartomeus"
date: "12/1/2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data collected in 2015 across 16 sites. Each site was visited 7 times. Each time we recorded three data types:

- All interactions in 100 m transects (`transects`)
- Focal interactions to selected plants (`focal`)
- Oportunistic new interactions not detected by the last two methods (`out`)

there are three datasets so far:

__Interactions:__ This include `transect`, `visitation` and `out` data. Data can be used pooled (1469 interactions ; 5706 visits) for general question abouts trait matching or splited by site (`Site_ID`) and `Round`.

```{r}
int <- read.csv("interactions.csv", h = T)
head(int)
nrow(int)
sum(int$Frequency, na.rm = TRUE)
```
Challenges:  
- long tail of singletones  
- not all species are taxonomically resolved. What to do with specimens we ony know Genus?
As for now:
  . sp means we only know the genus
  . morpho1,2,3 Unidentified species, but we know is a unique species diferent from anything else.
  . Family NA- means we only know the family. 
- Low sample size when you split per site and round

__pollinator traits:__
```{r}
pol_tr <- read.csv("traits_pollinators.csv", h = T)
head(pol_tr)
```
__NOTE:__ Most IT (intertegular length) for bees are actual measures. For floral visitors without wings we measured torax width. When species is unknown (Genus sp.) we calculated the mean of the genus or family. Tongue length is measured only for a few individuals. Bees tongue length for most species is estimated following Cariveau et al 2016. For families without an actual tongue the value is set to zero. For other species, tongue length is estimated to covary with allometry using same equation as for colletidaes (the shortest tongue bee family).

__plant traits:__
```{r}
plant_tr <- read.csv("traits_plants.csv", h = T)
head(plant_tr)
```
Those are complete.

Challenges:  
- We are not sure which trait are driving interactions (quite generalized system)

__independent plant flower abundance__: Ca be combined by Site_ID and Round
```{r}
flw <- read.csv("flowers.csv", h = T)
head(flw)
```

__Phylogenies__ Need to do.

```{r}
#Nacho this is the final list of species
pol_tr$Pollinator_gen_sp #277 entries. 
#sp means we don't know the species. Ignore for phylogeny and use a random species in this genus
#morpho1,2,3 Unidentified, but we know is a unique species. Idem. Ignore for phylogenies and use a species in genus as replacement?
#Family NA- means we only know the family. It will be intersting to make sure those families are represented, e.g. by adding a ranom species?
plant_tr$Plant_gen_sp
```

- Plants: Phylomatic?
- Pollinators: Bees available at genus level only. For other orders I have no idea.

# combining data

```{r}
head(int)
unique(int[,3:4])
unique(int$Plant_gen_sp[which(!int$Plant_gen_sp %in% plant_tr$Plant_gen_sp)]) #ok
unique(int$Pollinator_gen_sp[which(!int$Pollinator_gen_sp %in% pol_tr$Pollinator_gen_sp)])
temp <- merge(int, pol_tr, by.x = "Pollinator_gen_sp", by.y = "Pollinator_gen_sp", all.x = TRUE)
temp2 <- merge(temp, plant_tr, by.x = "Plant_gen_sp", by.y = "Plant_gen_sp", all.x = TRUE)
pooled_data <- temp2
head(pooled_data, 20)
```

# Reshaping the dataset as input of the `as.alienData()` function

**Extract from the roxygen2 documentation:**

```
#' @param idObs A data.frame which is mandatory and will help to check
consistency and prevent errors among unique identifiers of each alienData
arguments. The first column (idSite) contains unique identifier of where the
observation was made. The second column (idTime) is not mandatory and contains
temporal information: an unique identifier at the time the sample has been taken
(needed for timeseries analysis). The third column (idSpcies) is an unique
identifier of the species sampled at time (idTime) and location (idSite). The
fourth column is an unique identifier of individu of species (idSp) observed at
time (idTime) and location (idSite). 
#' @param interactPair A data.frame which
contains interaction at the finest level (individus or species). The first two
columns are idFrom and idTo and determine the sens of the interaction. idFrom
and idTo are unique identifier of species or individu documented in the idObs
data.frame. Finaly, the thrid column is the strength of the interaction (Please
see details). 
#' @param coOcc A square symmetric matrix of 0s and 1s that define
co-occurence patterns among pairs of species. If this matrix is not provided,
the co-occurence matrix is derived from the coAbund matrix else the interactSp
matrix (see return section). 
#' @param coAbund A square symmetric matrix that
includes any types of values, defining co-abundance patterns among pairs of
species. TODO: Not implemented yet. 
#' @param siteEnv A matrix or a data.frame
where each column is a descriptor of the sites. TODO: siteEnv should cover the
possibility that environmental variables could be taken at several times. 
#' @param traitSp A matrix or a data.frame where each column is a trait
characterizing all species. The first column is a unique identifier of the
species documented in idObs data.frame. 
#' @param traitInd A matrix or a data.frame where each column is a trait characterizing an individual. The first
column is a unique identifier of the individu documented in idObs data.frame. 
#' @param phylo A square symmetric matrix describing the phylogenetic relationships
between pairs of all species (see details). TODO: Not implemented yet. 
#' @param scaleSiteEnv Logical. Whether the columns of X should be centred and divided by
the standard deviation. Default is TRUE. 
#' @param scaleTrait Logical. Whether the rows of Tr should be centred and divided by the standard deviation. Default
is TRUE. 
#' @param interceptSiteEnv Logical. Whether a column of 1s should be
added to X. Default is TRUE. 
#' @param interceptTrait Logical. Whether a row of
1s should be added to Tr. Default is TRUE. 
#' @details 
#'The strength of the interactions defined in the third column of \code{interactPair} can be a 0
if no direct interaction has been observed (defined as true absence of
interaction) or any numerical value. Undocumented interactions among species or
individus will be assumed as NA by default.
```

```{r, echo=FALSE}
# Create dir for formatted data
#dir.create('./Data/alienData', showWarnings = FALSE)
```

### Each section below transform your data into as.alienData args

#### Get idObs

```{r}
# load data
int <- read.csv("interactions.csv", h = T, stringsAsFactors=FALSE)
flw <- read.csv("flowers.csv", h = T, stringsAsFactors=FALSE)
plant_tr <- read.csv("traits_plants.csv", h = T, stringsAsFactors=FALSE)
pol_tr <- read.csv("traits_pollinators.csv", h = T, stringsAsFactors=FALSE)

# Create unique IDs for pollinator individus
int$idPol <- seq(1,nrow(int),1) 
# Create unique IDs for plant individus
int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1) #technically, some individuus will be repeated, although there is not way to know. Good for now

# subset pollinator and plant with their own ids
pol <- unique(int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")]) #The unique is unececsary here, right?
colnames(pol) <- c("idSite","idTime","idSp","idInd")
plant <- unique(int[,c("Site_ID", "Round","Plant_gen_sp","idPlant")]) #idem
colnames(plant) <- c("idSite","idTime","idSp","idInd")

# Set idObs by merging both dfs
idObs <- rbind(pol,plant)
#head(idObs)

#Turn all columns as factor
idObs <- as.data.frame(lapply(idObs,as.factor))
```

#### Get InteracPair

```{r}
interactPair <- int[,c("idPol","idPlant","Frequency")]
```

#### traitInd

Trait by individus

```{r}
traitInd <- int[,c("idPol","Pollinator_sex")] 
traitInd$traitName <- "sex"
traitInd <- traitInd[,c('idPol','traitName','Pollinator_sex')]
names(traitInd)[3] <- 'value'
#head(traitInd)
```

#### traitSp

Trait by species
Note Removed family as a trait

```{r}
library(reshape2)
traitSpPlant <- melt(plant_tr[,-2],id=c("Plant_gen_sp"),value.name="value",variable.name="traitName")
names(traitSpPlant)[1] <- "idSp"

#head(pol_tr)
traitSpPol <- melt(pol_tr[,c(1,2,5)],id=c("Pollinator_gen_sp"),value.name="value",variable.name="traitName") #I also remove order
names(traitSpPol)[1] <- "idSp"

traitSp <- rbind(traitSpPlant,traitSpPol)
traitSp$idSp <- as.factor(traitSp$idSp)

#### errors
# Some entries in traitSp (merge of pol_tr and plant_tr) are documenting species which are not present in `ìnt`, the interaction pairwise data.frame.
#IB: This is ok! I collect traits for some species we do not have interactions.
removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]
traitSp <- subset(traitSp, !idSp %in% removeSpecies )
traitSp$idSp <- droplevels(traitSp$idSp)
```

#### Save all args of the as.alienData() into a RData file.

```{r}
save(idObs,interactPair,traitSp,traitInd,file="./alienData/argsAlienData.RData")
```