### 2017-02-15 Data/phylogenies.Rmd
Assembling phylogenies for empirical data

---
title: "Obtaining and assembling phylogenies"
author: "Ignacio Morales-Castilla"
date: "February 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading data
```{r reading data in}
## remove objects
rm(list=ls())
options(stringsAsFactors = FALSE)

setwd("~/GitHub/Montreal_Workshop/Data")
int <- read.csv("interactions.csv", h = T)
#head(int)

plants<-sub(" ","_",unique(int$Plant_gen_sp))
insects<-sub(" ","_",unique(int$Pollinator_gen_sp))

```

## Loading plant phylogeny and matching taxonomy

Out of the initial list of 56 plant species, 33 are in the phylogeny. For 21 of the remaining 23 species, we can assign the phylogenetic placement of a sister species (within the same genus). Two species can't be placed in this preliminary phylogeny (Mirtus_communis & Halimium_halimifolium). We could insert them as polytomies.

```{r phylo plants, echo=FALSE}
library(phytools)
plant.phylo<-read.tree("Plants_phylo.tre")

# plants not in phylogeny
not.in.phylo<-plants[which(!plants%in%plant.phylo$tip.label)]

# check if genus of these plants is in phylogeny
genus.plants<-sapply(strsplit(plants, "_"), "[", 1)
genus.phylo<-sapply(strsplit(plant.phylo$tip.label, "_"), "[", 1)
genus.not.in.phylo<-genus.plants[which(!genus.plants%in%genus.phylo)]

# retrieve sister species for those species not in phylo
rep.each.genus<-table(genus.plants)
new.names<-array(NA,length(not.in.phylo))
for (i in 1:length(not.in.phylo)){
  #print(i)
  sps.i<-not.in.phylo[i]
  genus.i<-strsplit(sps.i, "_")[[1]][1]
  length(which(genus.phylo!=genus.i))

  if(!genus.i%in%genus.not.in.phylo){
  if(length(which(genus.phylo==genus.i))>1){
  phylo.genus.i<-drop.tip(plant.phylo,which(genus.phylo!=genus.i))
  phylo.genus.i<-drop.tip(phylo.genus.i,which(phylo.genus.i$tip.label%in%plants))

  #if(rep.each.genus[genus.i]<length(phylo.genus.i$tip.label)){
    new.names[i]<-sample(phylo.genus.i$tip.label,1)
  } else {
    new.names[i]<-plant.phylo$tip.label[which(genus.phylo==genus.i)]
  }
  } 
}

# check that there are not repeated species
#sum(which(new.names%in%plants))

# replace names
plants.new<-plants
plants.new[which(!plants%in%plant.phylo$tip.label)]<-new.names
plants.new[is.na(plants.new)]<-"Mirtus_communis"

```

## The plant phylogeny
Phylogenetic tree with 54 plant species 
```{r phylo plants, echo=FALSE}
# prune and obtain final plant phylogeny:
plant.phylo.int<-drop.tip(plant.phylo,which(!plant.phylo$tip.label%in%plants.new))

# compare and get vector with original names in same order as names in phylogeny
final.sps.list<-unique(plants.new[which(plants.new%in%plant.phylo$tip.label)])
names.compar<-data.frame(names.phylo=final.sps.list,order.phylo=as.numeric(match(final.sps.list,plant.phylo.int$tip.label)),names.original=plants[-c(34,53)])

# reordering to match phylogeny
names.compar<-names.compar[with(names.compar,order(order.phylo)),]

# re-assign names to phylogeny
plant.phylo.int$tip.label<-names.compar$names.original
plot(plant.phylo.int,cex=0.6)

```

## Repeating for Pollinator phylogeny
This one is going to be trickier. Only 16 sps of our list of arthropods are included in TTOL. The following genera are not in the phylogeny. Should we go to familial level?

```{r phylo insects, echo=FALSE}
all.phylo<-read.tree("TTOL_arthropods.nwk")

# pollinators not in phylogeny
not.in.phylo<-insects[which(!insects%in%all.phylo$tip.label)]

# check if genus of these plants is in phylogeny
genus.insects<-sapply(strsplit(insects, "_"), "[", 1)
genus.phylo<-sapply(strsplit(all.phylo$tip.label, "_"), "[", 1)
genus.not.in.phylo<-unique(genus.insects[which(!genus.insects%in%genus.phylo)])
genus.not.in.phylo

```