### 2017-07-24 Data/alien_reshaping.Rmd
alienData v2 - new structure after some corrections inherents to methods

# Reshaping the dataset as input of the `as.alienData()` function

### Each section below transform your data into as.alienData args

#### Get idObs

```{r}
# load data
int <- read.csv("interactions.csv", h = T, stringsAsFactors=FALSE)
flw <- read.csv("flowers.csv", h = T, stringsAsFactors=FALSE)
plant_tr <- read.csv("traits_plants.csv", h = T, stringsAsFactors=FALSE)
pol_tr <- read.csv("traits_pollinators.csv", h = T, stringsAsFactors=FALSE)

# Create unique IDs for pollinator individus
int$idPol <- seq(1,nrow(int),1)
# Create unique IDs for plant individus
int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1) #technically, some individuus will be repeated, although there is not way to know. Good for now

# subset pollinator and plant with their own ids
pol <- unique(int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")]) #The unique is unececsary here, right?
colnames(pol) <- c("idSite","idTime","idSp","idInd")
plant <- unique(int[,c("Site_ID", "Round","Plant_gen_sp","idPlant")]) #idem
colnames(plant) <- c("idSite","idTime","idSp","idInd")

# Set idObs by merging both dfs
idObs <- rbind(pol,plant)
#head(idObs)

#Turn all columns as factor
idObs <- as.data.frame(lapply(idObs,as.factor))
```

#### Get InteracPair

```{r}
interactPair <- int[,c("idPol","idPlant","Frequency")]
```

#### traitInd

Trait by individus

```{r}
traitInd <- int[,c("idPol","Pollinator_sex")]
traitInd$traitName <- "sex"
traitInd <- traitInd[,c('idPol','traitName','Pollinator_sex')]
names(traitInd)[3] <- 'value'
#head(traitInd)
```

#### traitSp

Trait by species
Note Removed family as a trait

```{r}
library(reshape2)
traitSpPlant <- melt(plant_tr[,-2],id=c("Plant_gen_sp"),value.name="value",variable.name="traitName")
names(traitSpPlant)[1] <- "idSp"

#head(pol_tr)
traitSpPol <- melt(pol_tr[,c(1,2,5)],id=c("Pollinator_gen_sp"),value.name="value",variable.name="traitName") #I also remove order
names(traitSpPol)[1] <- "idSp"

traitSp <- rbind(traitSpPlant,traitSpPol)
traitSp$idSp <- as.factor(traitSp$idSp)

#### errors
# Some entries in traitSp (merge of pol_tr and plant_tr) are documenting species which are not present in `ìnt`, the interaction pairwise data.frame.
#IB: This is ok! I collect traits for some species we do not have interactions.
removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]
traitSp <- subset(traitSp, !idSp %in% removeSpecies )
traitSp$idSp <- droplevels(traitSp$idSp)
```

#### Save all args of the as.alienData() into a RData file.

```{r}
save(idObs,interactPair,traitSp,traitInd,file="./alienData/argsAlienData.RData")
```