### 2017-07-26 Data/alien_reshaping.Rmd
Reshaping to alienData v2 / tracking taxo with NCBI and ITIS

# Reshaping the dataset as input of the `as.alienData()` function

### Each section below transform your data into as.alienData args

#### Get idObs

```{r}
# load data
int <- read.csv("interactions.csv", h = T, stringsAsFactors=FALSE)
flw <- read.csv("flowers.csv", h = T, stringsAsFactors=FALSE)
plant_tr <- read.csv("traits_plants.csv", h = T, stringsAsFactors=FALSE)
pol_tr <- read.csv("traits_pollinators.csv", h = T, stringsAsFactors=FALSE)

# Create unique IDs for pollinator individus
int$idPol <- seq(1,nrow(int),1)
# Create unique IDs for plant individus
int$idPlant <- seq(nrow(int)+1,nrow(int)+nrow(int),1) # IB: technically, some individus will be repeated, although there is not way to know. Good for now

# subset pollinator and plant with their own ids
# if no_captured set his own ids
# int$idPol <- int$Pollinator_id
# id <- 1
# for(i in 1:nrow(int)){
#   if(int[i,'idPol'] == 'no_captured'){
#     int[i,'idPol'] <- id
#     id <- id+1
#   }
# }

pol <- int[,c("Site_ID", "Round","Pollinator_gen_sp","idPol")] #The unique is unececsary here, right? SV: yes
colnames(pol) <- c("idSite","idTime","idSp","idInd")
plant <- unique(int[,c("Site_ID", "Round","Plant_gen_sp","idPlant")]) #idem
colnames(plant) <- c("idSite","idTime","idSp","idInd")

# Set idObs by merging both dfs
idObs <- rbind(pol,plant)
head(idObs)

```

#### traitInd

Trait by individus

```{r}
traitInd <- int[,c("idPol","Pollinator_sex")]
traitInd$traitName <- "sex"
traitInd <- traitInd[,c('idPol','traitName','Pollinator_sex')]
names(traitInd)[3] <- 'value'
#head(traitInd)
```

#### Get InteracPair

```{r}
interactPair <- int[,c("idPol","idPlant","Frequency")]
```

#### traitSp

Trait by species
Note Removed family as a trait

```{r}
library(reshape2)
traitSpPlant <- melt(plant_tr[,-2],id=c("Plant_gen_sp"),value.name="value",variable.name="traitName")
names(traitSpPlant)[1] <- "idSp"

#head(pol_tr)
traitSpPol <- melt(pol_tr[,c(1,2,5)],id=c("Pollinator_gen_sp"),value.name="value",variable.name="traitName") #I also remove order
names(traitSpPol)[1] <- "idSp"

traitSp <- rbind(traitSpPlant,traitSpPol)
traitSp$idSp <- as.factor(traitSp$idSp)

#### errors
# Some entries in traitSp (merge of pol_tr and plant_tr) are documenting species which are not present in `ìnt`, the interaction pairwise data.frame.
#IB: This is ok! I collect traits for some species we do not have interactions.
# removeSpecies <- levels(traitSp$idSp)[which(!levels(traitSp$idSp) %in% levels(idObs$idSp))]
# traitSp <- subset(traitSp, !idSp %in% removeSpecies )
# traitSp$idSp <- droplevels(traitSp$idSp)
```

#### Prepare each aguments for alienData v2

```{r}

#### STEP A: dfNodes

dfNodes <- idObs[,c("idInd","idSp")]

## 1. ADD TRAITS FOR SP

dfNodes <- merge(dfNodes,traitSp,by="idSp",all=TRUE)
dfNodes <- dcast(idSp + idInd ~ traitName, data=dfNodes)

## 2. ADD TRAITS FOR IND

traitInd <- traitInd[,c('idPol','value')]
names(traitInd)[2] <- "sex"
dfNodes <- merge(dfNodes,traitInd,by.x="idInd",by.y="idPol",all=TRUE)

## 3. ADD TAXO
library(stringr)
dfNodes$species <- word(str_trim(dfNodes$idSp),-1)
dfNodes$genus <-  word(str_trim(dfNodes$idSp),-2)

## retrieve higher taxonomy based on ITIS
library(taxize)

cleanup_taxo <- function(x){
  if(all(!is.na(x))){
    if(!is.null(x) & dim(x)[1] != 1){
      r <- data.frame(t(x$name))
      colnames(r) <- x$rank
      return(r)
    } else {
      return(NA)
    }
  } else {
    return(NA)
  }
  }

# resolved_itis <- classification(unique(str_trim(dfNodes$idSp)),db="itis")
# saveRDS(resolved_itis,file='./resolved_itis.rds')
resolved_itis <- readRDS('resolved_itis.rds')

res_itis <- lapply(resolved_itis,cleanup_taxo)

res_itis <- res_itis[!is.na(res_itis)]
res_itis <- plyr::rbind.fill(res_itis)
res_itis <- res_itis[,names(res_itis)[which(colSums(apply(res_itis,2,is.na))==0)]]

# removing genus
res_itis <- res_itis[,-which(names(res_itis)=='genus')]
dfNodes <- merge(dfNodes,res_itis,by.x="idSp",by.y="species",all.x=TRUE)

## retrieve higher MISSING taxonomy based on GENUS with NCBI

taxo_missing <- unique(subset(dfNodes,is.na(family))[,'genus'])
resolved_ncbi <- readRDS('./resolved_ncbi.rds')
# resolved_ncbi <- classification(str_trim(taxo_missing),db="ncbi")
# saveRDS(resolved_ncbi,file='./resolved_ncbi.rds')

resolved_ncbi <- resolved_ncbi[!is.na(resolved_ncbi)]
res_ncbi <- lapply(resolved_ncbi,cleanup_taxo)
res_ncbi <- plyr::rbind.fill(res_ncbi)
res_ncbi <- res_ncbi[,c('genus',names(res_ncbi)[which(colSums(apply(res_ncbi,2,is.na))==0)])]

# remove taxo for species with no genus
res_ncbi <- subset(res_ncbi,!is.na(genus))
res_ncbi <- res_ncbi[,-2]

# cast classification items as factors
cols <- names(dfNodes)[11:19]
for(c in cols) dfNodes[,c] <- as.character(dfNodes[,c])

# fill missing columns on match
for(i in 1:nrow(res_ncbi)){
  dfNodes[which(res_ncbi[i,'genus'] == dfNodes$genus),'kingdom'] <- as.character(res_ncbi[i,'kingdom'])
  dfNodes[which(res_ncbi[i,'genus'] == dfNodes$genus),'order'] <- as.character(res_ncbi[i,'order'])
  dfNodes[which(res_ncbi[i,'genus'] == dfNodes$genus),'family'] <- as.character(res_ncbi[i,'family'])
}

## TODO: STILL TAXO LEVELS MISSINGS

## Final reshaping
dfNodes <- dfNodes[,c(2:ncol(dfNodes))]
names(dfNodes)[1] <- "idNodes"

#### STEP B: dfEdges

dfEdges <- int[,c('Site_ID','idPol','idPlant','Frequency')]
names(dfEdges) <- c('idSite','idFrom','idTo','value')

#### STEP C: dfSites

dfSites <- unique(dfEdges[,'idSite'])

#### Step D: dfOcc

dfOcc <- unique(dfEdges[,c('idSite','idTo')])
names(dfOcc)[2] <- "idSp"
dfOccFrom <- unique(dfEdges[,c('idSite','idFrom')])
names(dfOccFrom)[2] <- "idSp"
dfOcc <- rbind(dfOcc,dfOccFrom)

```

### Transform to alienData object

```{r}
plantsPol <-  alienData(dfNodes=dfNodes,dfEdges=dfEdges,verbose=TRUE,taxo=c('kingdom','class','order','family','genus','species'),trait=c('flower_morphology','flower_form','nectar_tube_diameter_mm','nectar_tube_depth_mm','IT','tongue_length','sex'))

save(plantsPol,file='../../alien/data/plantsPol.rda')

```